---
title: 'Statistical Methods for Discrete Response, Time Series, and Panel Data (W271):
  Lab 3'
author: "Professor Jeffrey Yau"
date: "March 18, 2018"
geometry: margin = 1 cm
output:
  pdf_document: default
---
\fontsize{8}{9.25}
\selectfont

******************************************************

## Introduction

<INSERT NARRATIVE>

### Load packages set some formatting preferences
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set digit print output to max desired
options(digits=2)
# Clean up the workspace before we begin
rm(list = ls())
# Load libraries
library(car)
library(dplyr)
library(Hmisc)
library(forecast)
library(astsa)
library(xts)
library(vars)
```

```{r Pkg and Data Load, message = FALSE, warning=FALSE}
pkg <- c('knitr', 'Hmisc', 'ggcorrplot', 'car',
         'dplyr', 'ggplot2', 'jtools', 'readr')
invisible(lapply(pkg, require, character.only = T))
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

### Question 1: E-Commerce Retail Sales as a Percent of Total Sales- Build a Seasonal ARIMA model and generate quarterly forecast for 2017

### Load data and quality-check the raw data
```{r}
# Load csv file as df
df_q1 <- read.csv("ECOMPCTNSA.csv", header = TRUE, sep=",")
str(df_q1)
#View(df_q1)
#names(df_q1)
head(df_q1)
describe(df_q1)
# Create an R time-series object with our quarterly data starting with final quarter of 1999
q1_ts <- ts(df_q1$ECOMPCTNSA, frequency = 4, start = c(1999,4))
str(q1_ts)
head(q1_ts)
```

We see that there are no missing values and that we are working with quarterly time series data.  No anomalies are detected and there is not potential for top or bottom code. On gross visual inspection of the time series, we see that the starting values are all less than 1, with the later values all being greater than 7.  This leads us to already suspect we are not dealing with a stationary series.  We cannot make a confident comment on seasonality without further EDA for visualization.

### EDA

#### Plot the data
```{r, out.width='.49\\linewidth', fig.width=7, fig.height=4, fig.show='hold', fig.align='center', strip.white=TRUE}
# For modeling purposes we keep data between 1999 and 2015 as our training data; we hold-out 2016 as test data
q1_ts_train <- q1_ts[time(q1_ts) >= 1999 & time(q1_ts) < 2016]
q1_ts_train <- ts(q1_ts_train, frequency = 4, start = c(1999,4))
q1_ts_test <- q1_ts[time(q1_ts) >= 2016]
q1_ts_test <- ts(q1_ts_test, frequency = 12, start = c(2016,1))
# Plot the training data
plot.ts(q1_ts_train, main="Quarterly data of E-Commerce Retail Sales as a Percent of Total Sales", ylab="Percent of total sales", col='blue')
```

We are now able to clearly see that the e-commernce retail sales time series is not stationary in the mean and exhibits seasonality. We will attempt to stationarize our time series via differencing.

#### Examine the ACF/PACF to determine if an AR(p) or MA(q) model is appropriate
```{r, out.width='.49\\linewidth', fig.width=7, fig.height=3.5, fig.show='hold', fig.align='center', strip.white=TRUE}
# Plot frequency distribution, ACF, PACF
hist(q1_ts_train, main="Frequency Distribution of E-commerce as Percentage of Total Sales", xlab="E-commerce as percentage of total sales")
acf(q1_ts_train, main="Autocorrelation function", lag.max = 24)
pacf(q1_ts_train, main="Partial Autocorrelation function", lag.max = 24)
```

Although it has been dictated that we are to create a SARIMA model, we still need to check that this model is appropriate.We see that the autocorrelations are significant for a large number of lag (16 quarters). This slow decay in the autocorrelations is evidence of a trend in the data, thus telling us that the time series is not stationary. We see that the partial autocorrelation plot has a significant spike at a lag of 1 quarter and at 4 quarters.  We see that our PACF has a somewhat abrupt drop-off non-seasonally following lag-1 and but that there are spikes in the PACF at an annual lag (multiples of lag-4 given our quarterly data).  This leads us to believe that our model will not require a non-seasonal MA term q but will require a seasonal MA term Q.  The histogram of our data shows that the e-commerce as percentage of total sales is fairly normally distributed withpositive skew; however, this tells us nothing about how the data are related in time.

#### Difference the data to impose stationarity



### Order Identification

### Model Creation: Build a Seasonal ARIMA model and generate quarterly forecast for 2017

### Fit Evaluation

### Question 2: data_2018Spring_MTS.txt

### Load data
```{r, out.width='.49\\linewidth', fig.width=5, fig.height=4, fig.show='hold', fig.align='center', strip.white=TRUE}
#df_q2 <- read.csv("correlate-flight_prices.csv", header = TRUE, sep=",")
```

### EDA

### Order Identification

### Model Creation

### Fit Evaluation

## Conclusion